---
title: "Retro studies"
author: "Yuelin Long"
date: "February 24, 2019"
output: html_document
---
```{r}
library(tidyverse)
```


#Problem 1
```{r}
#Input data
x1 = c(seq(25,75,10),seq(25,75,10))   #Set x1 as age. Use lower limit as the continous variable
x2 = c(rep(0,length(x1)/2),rep(1,length(x1)/2)) #Assume Daily Alcohol Consumption Control Daily Alcohol Consumption 0-79 g to be 0
y = c(0,5,21,34,36,8,1,4,25,42,19,5)
m_y = c(106,164,138,139,88,31,9,26,29,27,18,0)
m = y + m_y
resp = cbind(y, m_y)

#Build prospective model for retrospective study
logit.prosp=glm(resp~x1+x2, family=binomial(link='logit'))
summary(logit.prosp)
```

###Question 1

Intepretation:

+ The coefficient of x1 (b1): Given the daily alcohol consumption fixed, 1 year increase in age would lead to the log odds of getting esophageal cancer to increase by 0.061579;
+ The coefficient of x2 (b2): Given the age fixed, the log odds ratio of daily alcohol consumption of more than 80 g would be 1.780000 between case and control groups;
+ The intercept does not have any meaning because this is a retrospective study under a prospective model.



#Problem 2
```{r}
#Input data
x1 = c(rep(0,11),rep(1,10))  # Set x1 to be the species of seed. O. aegyptiaca 75 as 0 and O. aegyptiaca 73 as 1
x2 = c(rep(1,5),rep(0,6),rep(1,5),rep(0,5))  # Set x2 to be the root extract media (cucumber or bean). Bean as 1 and cucumber as 0.
y = c(10,23,23,26,17,5,53,55,32,46,10,8,10,8,23,0,3,22,15,32,3)
m = c(39,62,81,51,39,6,74,72,51,79,13,16,30,28,45,4,12,41,30,51,7)
resp = cbind(y,m - y)

#Build logistic regression model
logit_p2=glm(resp~x1+x2, family=binomial(link='logit'))
summary(logit_p2)
```

###Question 1

Intepretation:

+ The coefficient of x1 (b1): Given the root extract media fixed,  for aegyptiaca 73 seed, the log odds ratio between germination and non-germination seeds is -0.2705;
+ The coefficient of x2 (b2): Given the seed species fixed, for bean root extract media, the log odds ratio between germination and non-germination seeds is -1.0647;
+ The intercept: For aegyptiaca 75 seed and cucumber root extract media, the log odds ratio between germination and non-germination seeds is 0.6347;


###Question 2
```{r}
#Goodness of fit

#deviance
res_dev = logit_p2 %>% 
  broom::glance() %>% 
  select(deviance) %>% 
  as.numeric()
res_dev


#Generalized Pearson chi square statistic
res_pea = sum(residuals(logit_p2, type = "pearson")^2)
res_pea

#Critical value
n = 21
p = 3
qchisq(0.05,n-p)

pval=1-pchisq(res_dev,df=n-p)
pval
```

Both deviance and generalized pearson chi square statistics are far greater than critical value. Therefore, reject the null hypothesis and conclude that the fit is bad. The small p value also confirms it.

Use Half-normal plot to see whether it's overdispersion:
```{r}

phi=res_pea/(n-p)

plot(qnorm((n+1:n+0.5)/(2*n+1.125)),sort(abs(residuals(logit_p2, type = "pearson"))),xlab='Expected Half-Normal Order Stats',ylab='Ordered Abs Pearson Residuals')
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2)

phi
```

As we can see in the plot, the slope of our model deviates from the standard line (y=x), therefore there exists over dispersion. And the dispersion parameter is `r phi`.

Refit the model:
```{r}
summary(logit_p2,dispersion=phi)
```

The intepretation remains unchanged, but the standard of error increased.

Intepretation (same as before):

+ The coefficient of x1 (b1): Given the root extract media fixed,  for aegyptiaca 73 seed, the log odds ratio between germination and non-germination seeds is -0.2705;
+ The coefficient of x2 (b2): Given the seed species fixed, for bean root extract media, the log odds ratio between germination and non-germination seeds is -1.0647;
+ The intercept: For aegyptiaca 75 seed and cucumber root extract media, the log odds ratio between germination and non-germination seeds is 0.6347;


###Question 3
Over dispersion may caused by intra-class correlation or hierarchical sampling. The correlation between different group is highly unlikely, therefore the over dispersion should be caused by hierarchical sampling.
